Magnetised Polish doughnuts revisited

### 2017-06-28 paper_revised/reply_referee.tex

\bigskip

{\bf R:} - Sec. 2.1: they introduce the equation of state (13) but do not discuss
how they treat the specific internal energy that is part of the
enthalpy used in (13). A comment should be added.

\bigskip

{\bf A:} The specific internal energy $\varepsilon$ is obtained by combining the ideal fluid (gamma-law)
equation of state which, in the notation used in the paper we write $P=(\kappa-1)\rho\varepsilon$,
and the barotropic equation of state. Solving for $\varepsilon$ we obtain
\begin{eqnarray*}
\varepsilon=\frac{K}{\kappa-1}\rho^{\kappa-1}\,.
\end{eqnarray*}
We have incorporated the following statements at the end of the paragraph after Eq.~(14):
``We note that the specific internal energy $\varepsilon$ is needed to compute the enthalpy. We obtain
this quantity by combining the ideal fluid and the barotropic equations of state, which yields 
$\varepsilon=\frac{K}{\kappa-1}\rho^{\kappa-1}$."

{\bf Toni: creo que la expresi\'on anterior es incorrecta porque estoy usando $\rho$ cuando deber\'{\i}a usar $\omega$. Tenemos que
hablarlo.\}

### 2017-06-28 paper_revised/reply_referee.tex

\bigskip

{\bf R:} - Sec. 2.1: they introduce the equation of state (13) but do not discuss
how they treat the specific internal energy that is part of the
enthalpy used in (13). A comment should be added.

\bigskip

{\bf A:} We address this issue below where the referee comments on the ``assumption of the specific
internal energy", since we believe both queries are related.

@@

\bigskip

{\bf R:} - Sec. 3: the authors discuss that the rest-mass density distribution can
``be trivially obtained inverting the barotropic relation, $p = K
\rho^{\kappa}$". I am not sure this is so trivial unless an assumption is
made on the specific internal energy; the authors should discuss more
clearly how they actually obtain the density from Eqs. (12) and (14).

\bigskip

{\bf A:} 
The specific internal energy $\varepsilon$ is obtained by combining the ideal fluid (gamma-law)
equation of state which, in the notation used in the paper we write $P=(\kappa-1)\rho\varepsilon$,
and the barotropic equation of state. Solving for $\varepsilon$ we obtain
\begin{eqnarray*}
\varepsilon=\frac{K}{\kappa-1}\rho^{\kappa-1}\,.
\end{eqnarray*}
We have incorporated the following statements at the end of the paragraph after Eq.~(27):
``We note that the specific internal energy $\varepsilon$ is needed to compute the enthalpy. We obtain
this quantity by combining the ideal fluid and the barotropic equations of state, which yields 
$\varepsilon=\frac{K}{\kappa-1}\rho^{\kappa-1}$."

### 2017-06-28 paper_revised/reply_referee.tex

\bigskip

{\bf A:} 
The specific internal energy $\varepsilon$ is obtained by combining the ideal fluid (gamma-law)
equation of state which, in the notation used in the paper we write $P=(\kappa-1)\rho\varepsilon$,
and the barotropic equation of state. Solving for $\varepsilon$ we obtain
\begin{eqnarray*}
\varepsilon=\frac{K}{\kappa-1}\rho^{\kappa-1}\,.
\end{eqnarray*}
We have incorporated the following statements at the end of the paragraph after Eq.~(27):
``We note that the specific internal energy $\varepsilon$ is needed to compute the enthalpy. We obtain
this quantity by combining the ideal fluid and the barotropic equations of state, which yields 
$\varepsilon=\frac{K}{\kappa-1}\rho^{\kappa-1}$."

\bigskip

{\bf R:} - Sec. 3: the authors discuss that Eq.~(24) diverges for $\theta \rightarrow \pi/2$,
i.e. at the equator. I am a bit puzzled by this statement and I wonder if
this is an exact statement, ie whether it is true that

@@

\begin{equation}
\partial_{\theta} W = 0 \ \mathrm{for} \ \theta \rightarrow \frac{\pi}{2}.
\end{equation}

If this is an exact result, I urge the authors to prove it. If it is
not, I urge the authors to simply mention that the gradients become
very small.

\bigskip

{\bf A:} We believe this is an exact result and present the proof below:

Eq.~(22) of the paper gives us

\begin{eqnarray*}
\partial_{\theta} W = \partial_{\theta} \ln|u_t| - \frac{\Omega \partial_{\theta}l}{1 - \Omega l}\, = 0.
\end{eqnarray*}

Then, we have to prove that either the two terms of this equation are equal or they  are identically zero at $\theta = \pi/2$. We shall see that the latter is true.
First, we take the derivative of the angular momentum distribution $\partial_{\theta} l$. From Eq.~(7) we can write the angular momentum distribution as $l(r, \theta) = l(r) \sin^{2\gamma} \theta$, so that the derivative reads
\begin{eqnarray*}
\partial_{\theta} l = 2\gamma l(r) \sin^{2\gamma - 1} \theta \cos \theta \,.
\end{eqnarray*}
Taking $\theta = \pi/2$ leads to $\partial_{\theta} l = 0$, so the second term equals to zero.
To show that the first term is also zero, we write it as
\begin{eqnarray*}
\partial_{\theta} \ln|u_t| = \partial_{\theta} \ln \left(\frac{\mathcal{L}}{\mathcal{A}}\right)^{\frac{1}{2}}\,,
\end{eqnarray*}
where $\mathcal{L} = g_{t \phi}^2 - g_{tt} g_{\phi\phi}$ and $\mathcal{A} = g_{\phi\phi} + 2 l g_{t\phi} + l^2g_{tt}$, and we have dropped the absolute value, as it is irrelevant to this discussion. Then, the derivative is
\begin{eqnarray*}
\partial_{\theta} \ln|u_t| = \frac{1}{2} \frac{\mathcal{A}}{\mathcal{L}}\frac{\partial_{\theta}{\mathcal{L}}\mathcal{A} - \mathcal{L}\partial_{\theta}\mathcal{A}}{\mathcal{A}^2} = \frac{1}{2} \frac{\partial_{\theta}{\mathcal{L}}\mathcal{A} - \mathcal{L}\partial_{\theta}\mathcal{A}}{\mathcal{A} \mathcal{L}}\,.
\end{eqnarray*}
We use Boyer-Lindquist coordinates, for which the metric components read
\begin{eqnarray*}
g_{tt} = - \left(1 - \frac{2Mr}{\rho^2}\right), \ g_{t\phi} = -\frac{2Mar\sin^2\theta}{\rho^2}, \ g_{\phi\phi} = \left(r^2 + a^2 + \frac{2Ma^2r\sin^2\theta}{\rho^2}\right) \sin^2 \theta\,,
\end{eqnarray*}
where $\rho^2 = r^2 + a^2\cos^2 \theta$. Therefore, we obtain $\mathcal{L} = \Delta \sin^2 \theta$, where $\Delta = r^2 - 2Mr + a^2$, and its derivative $\partial_{\theta}\mathcal{L} = 2 \Delta \sin \theta \cos \theta$, which is zero for $\theta = \pi/2$. Next, we take the derivative of $\mathcal{A}$
\begin{eqnarray*}
\partial_{\theta} \mathcal{A} = \partial_{\theta} g_{\phi\phi} + 2 (\partial_{\theta} l) g_{t\phi} + 2l\partial_{\theta} g_{t\phi} + 2l(\partial_{\theta} l) g_{tt} + l^2 \partial_{\theta} g_{tt} = \partial_{\theta} g_{\phi\phi} + 2l\partial_{\theta} g_{t\phi} + l^2 \partial_{\theta} g_{tt}\,,
\end{eqnarray*}
where we have used the previous result that $\partial_{\theta} l=0$ at $\theta = \pi/2$. By inspecting the metric components, it is easy to see that all terms depending on $\theta$ are functions of $\sin^2 \theta$ or $\cos^2 \theta$. This means their $\theta$ derivatives will have at least a $\cos \theta$ multiplying. Then $\partial_{\theta} \mathcal{A} = 0$ at $\theta = \pi/2$. Therefore, $\partial_{\theta} \ln|u_t|$ is also zero at $\theta = \pi/2$.

We have incorporated this simple demonstration as a short appendix in the manuscript.

### 2017-06-28 paper_revised/reply_referee.tex

\bigskip

{\bf R:} - Fig. 1: as with the middle panel, I suggest that the right
panel should also show the logarithm of the rest-mass density.

\bigskip

{\bf A:} The reason we show the middle panel in logarithm scale and the right panel in linear scale is because this is the choice 
 in Komissarov (2006). We have followed the same choice in order to facilitate the qualitative and quantitative comparison
for this constant angular momentum test. If the referee concurred, we would prefer to keep the right panel in linear scale. If not, we would
not mind changing the right panel to logarithmic scale.

@@

First discussion of the dynamics of magnetized tori built with the Komissarov solution

{\tt http://adsabs.harvard.edu/abs/2007MNRAS.378.1101M}

\bigskip

{\bf A:} We have incorporated in the revised version of the article the first and the third references 
suggested by the referee. However, we have not incorporated the second 
reference. While this work is relevant in the context of oscillations 
of non-constant-angular momentum disks, including this reference in our 
work seems to us a bit forced since it discusses aspects which are a bit 
away from the main topic under discussion in our paper.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  
%\begin{small}
%\begin{thebibliography}{99}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  
%%
%
%%  
%\end{thebibliography}

% \end{small}

### 2017-06-28 paper_revised/paper_revised.tex
work on the paper
more work
first version of the appendices

Next, we choose all the initial radial values for the integration of Eq.~\eqref{eq:F} to lie between $r_{\mathrm{cusp}}$ and $r_{\mathrm{c}}$ ($\theta = \pi / 2$). Since we are only interested in the equipotential surfaces inside the Roche lobe of the disc, our choice of initial values provides us a mapping of the equipotential surfaces of the torus. Given that we have already obtained both the equipotential surfaces $\theta(r)$ which cross the segment $[r_{\mathrm{cusp}}, r_{\mathrm{c}}]$ at the equatorial plane and the values of the potential in that segment, we can obtain the complete potential distribution for the torus (outside of that segment). Once we have the potential distribution, we can find the {\bf fluid} pressure at the centre of the disc from Eq.~\eqref{eq:enthalpy_eq},
\begin{equation}
p_{\mathrm{c}} = w_{\mathrm{c}}(W_{\mathrm{in}} - W_{\mathrm{c}})\left(\frac{\kappa}{\kappa - 1} + \frac{\lambda
}{\lambda
 - 1}\frac{1}{\beta_{\mathrm{m}_{\mathrm{c}}}}\right)^{-1},
\end{equation}
where $w_{\mathrm{c}}$ is the enthalpy at the centre and
\begin{equation}
\label{eq:beta_eq}
\beta_{\mathrm{m}} = p/p_{\mathrm{m}},
\end{equation}
is the magnetisation parameter ($\beta_{\mathrm{m}_{\mathrm{c}}}$ being the magnetisation parameter at the centre of the disc). Using this definition, we can find the magnetic pressure at the centre,
\begin{equation}
p_{\mathrm{m_{\mathrm{c}}}} = p_{\mathrm{c}}/\beta_{\mathrm{m}_{\mathrm{c}}}\,.
\end{equation}
With both pressures known at the centre, fluid and magnetic, we can now find the constants $K$ and $K_{\mathrm{m}}$ using equations \eqref{eq:eos_fluid} and \eqref{eq:eos_mag}. Therefore, for a given inner radius of the disc $r_{\mathrm{in}}$ we can obtain the potential $W_{\mathrm{in}}$. We now finally have all the ingredients required to find the enthalpy distribution (Eq.~\eqref{eq:enthalpy_eq}), the fluid pressure and magnetic pressure distributions (Eqs.~\eqref{eq:eos_fluid} and \eqref{eq:eos_mag}), and the {\bf rest-mass} density distribution, which can be trivially obtained inverting the barotropic relation, $p = K \rho^{\kappa}$.

For the integration of Eq.~\eqref{eq:equatorial_pot} we use the composite Simpson's rule. It is important to use a very small integration step because the slope is very steep. In this work, we use a step $h = 10^{-6}$. Using the analytic, constant angular momentum case for comparison, we tested that 
a larger value of $h$ gives unacceptable accuracy losses. On the other hand, to integrate the ordinary differential equation \eqref{eq:F} we use a fourth-order Runge-Kutta method. Again, it is also important here to choose a suitable step of integration, especially at the outer end of the disc, because Eq.~\eqref{eq:F} diverges at the equatorial plane (the equipotential surfaces cross the equatorial plane perpendicularly).

@@

The dependence of the location of the maximum of the magnetic pressure with $\beta_{\mathrm{m}_{\mathrm{c}}}$ is shown in the right panel of Fig.~\ref{max-vs-magnetisation}. While this location also decreases with decreasing $\beta_{\mathrm{m}_{\mathrm{c}}}$, the constant value achieved for values above $\beta_{\mathrm{m}_{\mathrm{c}}}\sim 10^{3}$ does depend on the specific values of $\beta$ and $\gamma$, contrary to what happens with the fluid pressure. It must be noted that the location of the maximum of the magnetic pressure is nearly identical for all models considered when $\log_{10}\beta_{\mathrm{m}_{\mathrm{c}}} \sim 0.5$. At this value of $\beta_{\mathrm{m}_{\mathrm{c}}}$ all models seem to cross, as it is more clearly shown in the inset of the right panel of Fig.~\ref{max-vs-magnetisation}.

@@

>>>>

Next, we choose all the initial radial values for the integration of Eq.~\eqref{eq:F} to lie between $r_{\mathrm{cusp}}$ and $r_{\mathrm{c}}$ ($\theta = \pi / 2$). Since we are only interested in the equipotential surfaces inside the Roche lobe of the disc, our choice of initial values provides us a mapping of the equipotential surfaces of the torus. Given that we have already obtained both the equipotential surfaces $\theta(r)$ which cross the segment $[r_{\mathrm{cusp}}, r_{\mathrm{c}}]$ at the equatorial plane and the values of the potential in that segment, we can obtain the complete potential distribution for the torus (outside of that segment). Once we have the potential distribution, we can find the {\bf fluid} pressure at the centre of the disc from Eq.~\eqref{eq:enthalpy_eq},
\begin{equation}
p_{\mathrm{c}} = w_{\mathrm{c}}(W_{\mathrm{in}} - W_{\mathrm{c}})\left(\frac{\kappa}{\kappa - 1} + \frac{\lambda
}{\lambda
 - 1}\frac{1}{\beta_{\mathrm{m}_{\mathrm{c}}}}\right)^{-1},
\end{equation}
where $w_{\mathrm{c}}$ is the enthalpy at the centre and
\begin{equation}
\label{eq:beta_eq}
\beta_{\mathrm{m}} = p/p_{\mathrm{m}},
\end{equation}
is the magnetisation parameter ($\beta_{\mathrm{m}_{\mathrm{c}}}$ being the magnetisation parameter at the centre of the disc). Using this definition, we can find the magnetic pressure at the centre,
\begin{equation}
p_{\mathrm{m_{\mathrm{c}}}} = p_{\mathrm{c}}/\beta_{\mathrm{m}_{\mathrm{c}}}\,.
\end{equation}
With both pressures known at the centre, fluid and magnetic, we can now find the constants $K$ and $K_{\mathrm{m}}$ using equations \eqref{eq:eos_fluid} and \eqref{eq:eos_mag}. Therefore, for a given inner radius of the disc $r_{\mathrm{in}}$ we can obtain the potential $W_{\mathrm{in}}$. We now finally have all the ingredients required to find the enthalpy distribution (Eq.~\eqref{eq:enthalpy_eq}), the fluid pressure and magnetic pressure distributions (Eqs.~\eqref{eq:eos_fluid} and \eqref{eq:eos_mag}), and the {\bf rest-mass} density distribution, which can be trivially obtained inverting the barotropic relation, $p = K \rho^{\kappa}$. {\bf We note that the specific internal energy $\varepsilon$ is needed to compute the enthalpy. We obtain
this quantity by combining the ideal fluid and the barotropic equations of state, which yields 
$\varepsilon=\frac{K}{\kappa-1}\rho^{\kappa-1}$.}

For the integration of Eq.~\eqref{eq:equatorial_pot} we use the composite Simpson's rule. It is important to use a very small integration step because the slope is very steep. In this work, we use a step $h = 10^{-6}$. Using the analytic, constant angular momentum case for comparison, we tested that 
a larger value of $h$ gives unacceptable accuracy losses. On the other hand, to integrate the ordinary differential equation \eqref{eq:F} we use a fourth-order Runge-Kutta method. Again, it is also important here to choose a suitable step of integration, especially at the outer end of the disc, because Eq.~\eqref{eq:F} diverges at the equatorial plane (the equipotential surfaces cross the equatorial plane perpendicularly). {\bf We show the proof for this statment at appendix \ref{div_partial_W}.}

@@

The dependence of the location of the maximum of the magnetic pressure with $\beta_{\mathrm{m}_{\mathrm{c}}}$ is shown in the right panel of Fig.~\ref{max-vs-magnetisation}. While this location also decreases with decreasing $\beta_{\mathrm{m}_{\mathrm{c}}}$, the constant value achieved for values above $\beta_{\mathrm{m}_{\mathrm{c}}}\sim 10^{3}$ does depend on the specific values of $\beta$ and $\gamma$, contrary to what happens with the fluid pressure. {\bf It must be noted that the location of the maximum of the magnetic pressure is nearly identical for all models considered when $\beta_{\mathrm{m}_{\mathrm{c}}} = 1/(\lambda - 1) = 3$ ({\bf we show the proof for this affirmation at the appendix \ref{app_magmax})}. At this value of $\beta_{\mathrm{m}_{\mathrm{c}}}$ all models cross at $r_{\mathrm{m}_{\mathrm{max}}} = r_{\mathrm{c}}$, as it is more clearly shown in the inset of the right panel of Fig.~\ref{max-vs-magnetisation}}

% It must be noted that the location of the maximum of the magnetic pressure is nearly identical for all models considered when $\log_{10}\beta_{\mathrm{m}_{\mathrm{c}}} \sim 0.5$. At this value of $\beta_{\mathrm{m}_{\mathrm{c}}}$ all models seem to cross, as it is more clearly shown in the inset of the right panel of Fig.~\ref{max-vs-magnetisation}.

@@

\begin{appendix}
\section{Divergence of equation \eqref{eq:F} at $\theta = \pi/2$}\label{div_partial_W}

Eq.~\eqref{eq:polar_der_pot} of the paper gives us

\begin{eqnarray*}
\partial_{\theta} W = \partial_{\theta} \ln|u_t| - \frac{\Omega \partial_{\theta}l}{1 - \Omega l}\, = 0.
\end{eqnarray*}

Then, we have to prove that either the two terms of this equation are equal or they are identically zero at $\theta = \pi/2$. We shall see that the latter is true.
First, we take the derivative of the angular momentum distribution $\partial_{\theta} l$. From Eq.~(7) we can write the angular momentum distribution as $l(r, \theta) = l(r) \sin^{2\gamma} \theta$, so that the derivative reads
\begin{eqnarray*}
\partial_{\theta} l = 2\gamma l(r) \sin^{2\gamma - 1} \theta \cos \theta \,.
\end{eqnarray*}
Taking $\theta = \pi/2$ leads to $\partial_{\theta} l = 0$, so the second term equals to zero.
To show that the first term is also zero, we write it as
\begin{eqnarray*}
\partial_{\theta} \ln|u_t| = \partial_{\theta} \ln \left(\frac{\mathcal{L}}{\mathcal{A}}\right)^{\frac{1}{2}}\,,
\end{eqnarray*}
where $\mathcal{L} = g_{t \phi}^2 - g_{tt} g_{\phi\phi}$ and $\mathcal{A} = g_{\phi\phi} + 2 l g_{t\phi} + l^2g_{tt}$, and we have dropped the absolute value, as it is irrelevant to this discussion. Then, the derivative is
\begin{eqnarray*}
\partial_{\theta} \ln|u_t| = \frac{1}{2} \frac{\mathcal{A}}{\mathcal{L}}\frac{\mathcal{A}\partial_{\theta}{\mathcal{L}} - \mathcal{L}\partial_{\theta}\mathcal{A}}{\mathcal{A}^2} = \frac{1}{2} \frac{\mathcal{A}\partial_{\theta}{\mathcal{L}} - \mathcal{L}\partial_{\theta}\mathcal{A}}{\mathcal{A} \mathcal{L}}\,.
\end{eqnarray*}
We use Boyer-Lindquist coordinates, for which the metric components read
\begin{eqnarray*}
g_{tt} = - \left(1 - \frac{2Mr}{\rho^2}\right), \ g_{t\phi} = -\frac{2Mar\sin^2\theta}{\rho^2}, \ g_{\phi\phi} = \left(r^2 + a^2 + \frac{2Ma^2r\sin^2\theta}{\rho^2}\right) \sin^2 \theta\,,
\end{eqnarray*}
where $\rho^2 = r^2 + a^2\cos^2 \theta$. Therefore, we obtain $\mathcal{L} = \Delta \sin^2 \theta$, where $\Delta = r^2 - 2Mr + a^2$, and its derivative $\partial_{\theta}\mathcal{L} = 2 \Delta \sin \theta \cos \theta$, which is zero for $\theta = \pi/2$. Next, we take the derivative of $\mathcal{A}$
\begin{eqnarray*}
\partial_{\theta} \mathcal{A} = \partial_{\theta} g_{\phi\phi} + 2 (\partial_{\theta} l) g_{t\phi} + 2l\partial_{\theta} g_{t\phi} + 2l(\partial_{\theta} l) g_{tt} + l^2 \partial_{\theta} g_{tt} = \partial_{\theta} g_{\phi\phi} + 2l\partial_{\theta} g_{t\phi} + l^2 \partial_{\theta} g_{tt}\,,
\end{eqnarray*}
where we have used the previous result that $\partial_{\theta} l=0$ at $\theta = \pi/2$. By inspecting the metric components, it is easy to see that all terms depending on $\theta$ are functions of $\sin^2 \theta$ or $\cos^2 \theta$. This means their $\theta$ derivatives will have at least a $\cos \theta$ multiplying. Then $\partial_{\theta} \mathcal{A} = 0$ at $\theta = \pi/2$. Therefore, $\partial_{\theta} \ln|u_t|$ is also zero at $\theta = \pi/2$.

\section{Magnetization for $r_{\mathrm{m}_{\mathrm{max}}} = r_{\mathrm{c}}$}\label{app_magmax}

Here, we derive the condition $\beta_{\mathrm{m}_{\mathrm{c}}} = 1/(\lambda - 1)$ for $r_{\mathrm{m}_{\mathrm{max}}} = r_{\mathrm{c}}$.
First, we can use equation~\eqref{eq:eos_mag} to write

\begin{equation}\label{eq_max}
\frac{\partial{p_{\mathrm{m}}(r)}}{\partial r} = \partial_{r} \left(K_{\mathrm{m}} \mathcal{L}^{\lambda -1} w^{\lambda}\right) = 0,
\end{equation}

Computing, we arrive at

\begin{equation}
\partial_{r} p_{\mathrm{m}} = K_{\mathrm{m}} \mathcal{L}^{\lambda - 2}w^{\lambda} [(\lambda - 1)(\partial_r \mathcal{L}) w + \lambda \mathcal{L} (\partial_r w)].
\end{equation}

We can see that inside the disc $K_{\mathrm{m}} \mathcal{L}^{\lambda - 2}w^{\lambda} \neq 0$, So to fulfill the extremum condition \eqref{eq_max} we need

\begin{equation}\label{eq:simple_expr}
(\lambda - 1)(\partial_r \mathcal{L}) w + \lambda \mathcal{L} (\partial_r w) = 0
\end{equation}

To evaluate this expressión, we need to compute the partial derivatives $\partial_r \mathcal{L}$ and $\partial_r w$. The derivative of $\mathcal{L}$ is esay to compute, as $\mathcal{L} = \Delta \sin^2 \theta$. So its partial derivative is

\begin{equation}
\partial_r \mathcal{L} = 2(r-M)\sin^2 \theta
\end{equation}

Now, we take our attention to $\partial_r w$. From eq.~\eqref{eq:enthalpy_eq} we have

\begin{equation}
w = \left[-\Delta W \left(\frac{\lambda -1}{\lambda}\right)\frac{1}{K+K_{\mathrm{m}}\mathcal{L}^{\lambda - 1}}\right]^{\frac{1}{\lambda-1}}
\end{equation}

Then, we can take its derivative

\begin{equation}
\partial_r w = \frac{w}{\lambda - 1} \left[-\Delta W \left(\frac{\lambda -1}{\lambda}\right)\frac{1}{K+K_{\mathrm{m}}\mathcal{L}^{\lambda - 1}}\right]^{-1} \left[\left(-\partial_r W \left(\frac{\lambda -1 }{\lambda}\right) \frac{1}{K+K_{\mathrm{m}}\mathcal{L}^{\lambda - 1}}\right) + \left(-\Delta W \left(\frac{\lambda -1}{\lambda}\right)\frac{-(\lambda - 1) K_{\mathrm{m}} \mathcal{L}^{\lambda - 2} \partial_r \mathcal{L}}{(K+K_{\mathrm{m}}\mathcal{L}^{\lambda - 1})^2}\right)\right]
\end{equation}

Since we are at $r = r_{\mathrm{c}}$ the derivative of the potential is zero ($\partial_r W(r, \pi/2)|_{r = r_{\mathrm{c}}} = 0$), so we can simplify the expression to get

\begin{equation}
\partial_r w = \frac{w}{\lambda - 1} \left[-\Delta W \left(\frac{\lambda -1}{\lambda}\right)\frac{1}{K+K_{\mathrm{m}}\mathcal{L}^{\lambda - 1}}\right]^{-1} \left(-\Delta W \left(\frac{\lambda -1}{\lambda}\right)\frac{1}{K+K_{\mathrm{m}}\mathcal{L}^{\lambda - 1}}\right)\left[\frac{-(\lambda - 1) K_{\mathrm{m}} \mathcal{L}^{\lambda - 2} \partial_r \mathcal{L}}{K+K_{\mathrm{m}}\mathcal{L}^{\lambda - 1}}\right]
\end{equation}

And we can further simplify this, to arrive at

\begin{equation}
\partial_r w = - w \left[\frac{K_{\mathrm{m}} \mathcal{L}^{\lambda - 2} \partial_r \mathcal{L}}{K+K_{\mathrm{m}}\mathcal{L}^{\lambda - 1}}\right]
\end{equation}

Then, we can insert this expression at eq.~\eqref{eq:simple_expr} to obtain

\begin{equation}
(\partial_r \mathcal{L}) w [(\lambda -1)(K+K_{\mathrm{m}}\mathcal{L}^{\lambda - 1}) - \lambda K_{\mathrm{m}}\mathcal{L}^{\lambda - 1}] = 0
\end{equation}

As $(\partial_r \mathcal{L}) w \neq 0$ inside the disc, we can simplify this, to arrive at

\begin{equation}
K (\lambda - 1) = K_{\mathrm{m}}\mathcal{L}^{\lambda - 1}
\end{equation}

The, we can use equations \eqref{eq:eos_fluid}, \eqref{eq:eos_mag} and \eqref{eq:beta_eq} at the center (remember that $w_{\mathrm{c}} = 1$) to write $K$ as $K = \beta_{\mathrm{m}_{\mathrm{c}}} K_\mathrm{m}\mathcal{L}^{\lambda -1}$. This allows us to write the simple expression

\begin{equation}
\beta_{\mathrm{m}_{\mathrm{c}}} = \frac{1}{\lambda - 1}
\end{equation}

\sg{We should add a short discussion of the generality of this result.}

\end{appendix}
